---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# tidycomm

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/tidycomm)](https://CRAN.R-project.org/package=tidycomm)
<!-- badges: end -->

Tidycomm provides convenience functions for common tasks in communication research. Currently, tidycomm includes functions for various methods of univariate and bivariate data description and analysis, data modification, and intercoder reliability tests. All functions follow the style and syntax of the [tidyverse](https://www.tidyverse.org/).

## Installation

Install the most recent development version of tidycomm with:

```{r eval = FALSE}
devtools::install_github("joon-e/tidycomm")
```

