---
title: "Univariate analysis of continuous and categorical variables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{univariate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The first step in data exploration usually consists of univariate, descriptive analysis of all variables of interest. Tidycomm offers two basic functions to quickly output relevant statistics:

- `describe()` for continuous variables
- `tab_frequencies()` for categorical variables

```{r setup, message = FALSE, warning = FALSE}
library(tidycomm)
library(tidyverse)
```

## Describe continuous variables with `describe()`

`describe()` outputs several measures of central tendency and variability for all variables named in the function call:

```{r}
mtcars %>% 
  describe(mpg, disp, hp)
```

If no variables are passed to `describe()`, all numeric variables in the data are described:

```{r}
mtcars %>% 
  describe()
```

Data can be grouped before describing:

```{r}
mtcars %>% 
  group_by(cyl) %>% 
  describe(mpg, disp, hp)
```

## Tabulate frequencies of categorical variables with `tab_frequencies()`

`tab_frequencies()` outputs absolute and relative frequencies of all unique values of one or more categorical variables:

```{r}
diamonds %>% 
  tab_frequencies(clarity)
```

Passing more than one variable will compute relative frequencies based on all combinations of unique values:

```{r}
diamonds %>% 
  tab_frequencies(clarity, cut)
```

You can also group your data before. This will lead to within-group relative frequencies:

```{r}
diamonds %>% 
  group_by(cut) %>% 
  tab_frequencies(clarity)
```

(Compare the columns `percent` and `cum_percent` with the output above.)
