---
title: "Change Scales"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scales}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(rmarkdown.html_vignette.check_title = FALSE)
```

Tidycomm provides four functions to easily translate continuous scales and to standardize them:

- `reverse_scale()` simply turns a scale upside down
- `minmax_scale()` down- or upsizes a scale to new minimum/maximum while retaining distances
- `center_scale()` subtracts the mean from each individual data point to center a scale at a mean of 0
- `z_scale()` works just like `center_scale()` but also divides the result by the standard deviation to also obtain a standard deviation of 1 and make it comparable to other z-standardized distributions

These functions provide convenience wrappers that make it easy to read and spell out how you transformed your scales.

```{r setup}
library(tidycomm)
```
The easiest one is to reverse your scale. You can just specify the scale and define the scale's lower and upper end. Take `autonomy_emphasis` as an example that originally ranges from 1 to 5. We will reverse it to range from 5 to 1.

The function adds a new column named `autonomy_emphasis_rev`: 

```{r}
WoJ %>% 
  reverse_scale(autonomy_emphasis,
                lower_end = 1,
                upper_end = 5) %>% 
  dplyr::select(autonomy_emphasis,
                autonomy_emphasis_rev)
```

Alternatively, you can also specify the new column name manually:

```{r}
WoJ %>% 
  reverse_scale(autonomy_emphasis,
                name = "new_emphasis",
                lower_end = 1,
                upper_end = 5) %>% 
  dplyr::select(autonomy_emphasis,
                new_emphasis)
```

`minmax_scale()` just takes your continuous scale to a new range. For example, convert the 1-5 scale of `autonomy_emphasis` to a 1-10 scale while keeping the distances:

```{r}
WoJ %>% 
  minmax_scale(autonomy_emphasis,
               change_to_min = 1,
               change_to_max = 10) %>% 
  dplyr::select(autonomy_emphasis,
                autonomy_emphasis_1to10)
```

`center_scale()` moves your continuous scale around a mean of 0:

```{r}
WoJ %>% 
  center_scale(autonomy_selection) %>% 
  dplyr::select(autonomy_selection,
                autonomy_selection_centered)
```

Finally, `z_scale()` does more or less the same but standardizes the outcome. To visualize this, we look at it with a visualized `tab_frequencies()`:


```{r}
WoJ %>% 
  z_scale(autonomy_selection) %>% 
  tab_frequencies(autonomy_selection,
                  autonomy_selection_z) %>% 
  visualize()
```
